---
title: "Forecast_w_social_predictor_John"
author: "Zhang"
date: "2025-04-20"
output: pdf_document
---


----------

This code contains the prediction of monthly methane not using social variables.

With each model, it contains two parts:

`### Choose the best variable combination`: A code to find the best social variable combination
`### Model`: The fitted model using the best variable combination

In the code chunk `# Combination of results`, you can see all the predicted values, you can directly paste them for combination.

You may also open the outline of this document, it is more clear for you to get the structure of the code.

----------


```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)

set.seed(797)

```


# Preparation: Import and split data to training and test sets

```{r }
# Import social economic factors with methane
methane_all <- read.csv("./Data/Processed/social_economic_factors_monthly.csv")


# Transform to time series
methane_all_ts <- ts(methane_all, 
                start = c(1983,7), 
                end = c(2024,11),
                frequency = 12)

methane_all_msts <- msts(methane_all, 
                            seasonal.periods =c(4,12),
                            start = c(1983,7),
                            end = c(2024,11))


# Split to training and test sets
methane_train_ts <- window(methane_all_ts, end=c(2021,12))
methane_test_ts <- window(methane_all_ts, start=c(2022,1))

methane_train_msts <- window(methane_all_msts, end=c(2021,12))
methane_test_msts <- window(methane_all_msts, start=c(2022,1))



```


# Predictions


## Arima

### Choose the best variable combination

```{r echo=FALSE}

variables <- list(c(2),c(3),c(4),c(5))
result <- list()

for(variable in variables){
  
  #autofit the sarima model
arima_autofit <- auto.arima(methane_train_ts[,1], 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0,
                            xreg = methane_train_ts[,variable])


#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = 35,xreg = methane_test_ts[,variable])

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima$mean, methane_test_ts[,1])
  
  result_list <- data.frame(Variable_colnum = variable, RMSE = forecast_arima_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(forecast_arima_accuracy)


}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)
```


### Arima

```{r }
arima_autofit <- auto.arima(methane_train_ts[,1], 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0,
                            xreg = methane_train_ts[,5])


forecast_arima <- forecast(object = arima_autofit, h = 35,xreg = methane_test_ts[,5])


forecast_arima_accuracy <- accuracy(forecast_arima$mean, methane_test_ts[,1])

print(forecast_arima_accuracy)

plot(forecast_arima)

#write to csv
write.csv(forecast_arima$mean,
          "./Forecasts/Methane/arimaExoSE.csv",
          row.names = FALSE)
```


## ARIMA with Fourier

### Choose the best variable combination

```{r echo=FALSE}

#variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

xreg_train <- cbind(fourier(methane_train_msts[,1],K=c(2,6)), methane_train_msts[,variable])
xreg_test <- cbind(fourier(methane_test_msts[,1],K=c(2,6),h=35), methane_test_msts[,variable])


arima_fourier_autofit <- auto.arima(methane_train_msts[,1],
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=xreg_train)


#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg= xreg_test)
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts[,1])
  
  
  #####
  # Find the best variable
  #####
  
  result_list <- data.frame(Variable_colnum = variable, RMSE = forecast_arima_fourier_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(forecast_arima_fourier_accuracy)

}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```
### ARIMA with Fourier

```{r}
#autofit the arima model

xreg_train <- cbind(fourier(methane_train_msts[,1],K=c(2,6)), methane_train_msts[,2])
xreg_test <- cbind(fourier(methane_test_msts[,1],K=c(2,6),h=35), methane_test_msts[,2])


arima_fourier_autofit <- auto.arima(methane_train_msts[,1],
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=xreg_train)


#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg= xreg_test)
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts[,1])


print(forecast_arima_fourier_accuracy)

#write to csv
write.csv(forecast_arima_fourier$mean,
          "./Forecasts/Methane/arimaFourierExoSE.csv",
          row.names = FALSE)
```


## SARIMA

### Choose the best variable combination
```{r echo=FALSE}

#variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))
result <- list()

for(variable in variables){

sarima_autofit <- auto.arima(methane_train_ts[,1],xreg = methane_train_ts[,variable])

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = 35,xreg = methane_test_ts[,variable])

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts[,1])


  result_list <- data.frame(Variable_colnum = variable, RMSE = forecast_sarima_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(forecast_sarima_accuracy)
  
}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```


### SARIMA
```{r}
#autofit the sarima model
sarima_autofit <- auto.arima(methane_train_ts[,1],xreg = methane_train_ts[,5])

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = 35,xreg = methane_test_ts[,5])

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts[,1])

#plot the forecast
plot(forecast_sarima)

#write to csv
write.csv(forecast_sarima$mean,
          "./Forecasts/Methane/sarimaExoSE.csv",
          row.names = FALSE)
```



## TBATS 

### Choose the best variable combination
```{r echo=FALSE}

#variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

 #fit TBATS
tbats_fit <- tbats(methane_train_msts[,1],xreg = methane_train_msts[,variable])

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_msts[,1]),xreg = methane_test_msts[,variable])

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_msts[,1]) 
  


  
  result_list <- data.frame(Variable_colnum = variable, RMSE = tbats_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(tbats_accuracy)

}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print(best_result)

```
### TBAS

```{r}
#fit TBATS
tbats_fit <- tbats(methane_train_msts[,1],xreg = methane_train_msts[,2])

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_msts[,1]),xreg = methane_test_msts[,2])

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_msts[,1])

#visualization
plot(tbats_forecast)

print(tbats_accuracy)

#write to csv
write.csv(tbats_forecast$mean,
          "./Forecasts/Methane/tbatsExoSE.csv",
          row.names = FALSE)

```


## Neuron Network

### Choose the best variable combination


```{r echo=FALSE}

variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

#variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

NN_fit <- nnetar(methane_train_ts[,1], xreg = methane_train_ts[,variable], p = 0, P = 7, repeats = 100)

NN_for <- forecast(NN_fit, xreg = methane_test_ts[,variable], h=35)

forecast_NN_accuracy <- accuracy(NN_for$mean, methane_test_ts[,1])
  
  
  #####
  # Find the best variable
  #####
  
  result_list <- data.frame(
    Variable_colnum = paste(variable, collapse = ","),
    MAPE = forecast_NN_accuracy["Test set", "MAPE"]
  )
  
  result <- append(result, list(result_list))
  
  cat("Variables:", variable, "\n")
  print(forecast_NN_accuracy)

}

rmse_values <- sapply(result, function(x) x$MAPE)
best_index <- which.min(rmse_values)
best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```

### Neural Network

```{r}

NN_fit <- nnetar(methane_train_ts[,1], xreg = methane_train_ts[,c(2,4)], p = 0, P = 7,repeats = 100)

NN_for <- forecast(NN_fit, xreg = methane_test_ts[,c(2,4)], h=35)

forecast_NN_accuracy <- accuracy(NN_for$mean, methane_test_ts[,1])

print(forecast_NN_accuracy)

plot(NN_for)

#write to csv
write.csv(NN_for$mean,
          "./Forecasts/Methane/nnExoSE.csv",
          row.names = FALSE)
```


## STL+Arima
Note: STL only supports arima model for external variables

### Choose the best variable combination

```{r echo=FALSE}

#variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

stl_arima_fit <- stlm(methane_train_ts[,1], xreg = methane_train_ts[,variable], s.window = "periodic", method = "arima")

#forecasting test data
stl_arima_forecast <- forecast(stl_arima_fit, h = length(methane_test_ts), xreg = methane_test_ts[,variable])

#accuracy assessment
stl_arima_accuracy <- accuracy(stl_arima_forecast$mean, methane_test_ts[,1])
  

  cat(variable,":\n")
  
  print(stl_arima_accuracy)
  
  #####
  # Find the best variable
  #####
  
  result_list <- data.frame(Variable_colnum = variable, RMSE = stl_arima_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
 

}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```

### STL+Arima

```{r}
#fit STL with ETS
stl_arima_fit <- stlm(methane_train_ts[,1], xreg = methane_train_ts[,2], s.window = "periodic", method = "arima")

#forecasting test data
stl_arima_forecast <- forecast(stl_arima_fit, h = length(methane_test_ts), xreg = methane_test_ts[,2])

#accuracy assessment
stl_arima_accuracy <- accuracy(stl_arima_forecast$mean, methane_test_ts[,1])

#visualization
plot(stl_arima_forecast)

print(stl_arima_accuracy)

#write to csv
write.csv(stl_arima_forecast$mean,
          "./Forecasts/Methane/stlArimaExoSE.csv",
          row.names = FALSE)
```

## State Space


### Choose the best variable combination

```{r echo=FALSE}

#variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

SSES <- es(methane_train_ts[,1],model="ZZZ",h=35,holdout=FALSE, xreg = methane_train_ts[,variable])

SSES_for <-forecast(SSES,h=35, interval="prediction", xreg = methane_test_ts[,variable])

forecast_SSES_accuracy <- accuracy(SSES$forecast, methane_test_ts[,1])
  
  
  #####
  # Find the best variable
  #####
  
  result_list <- data.frame(Variable_colnum = variable, RMSE = forecast_SSES_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(forecast_SSES_accuracy)

}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```

### State Space

```{r}


SSES <- es(methane_train_ts[,1],model="ZZZ",h=35,holdout=FALSE, xreg = methane_train_ts[,5])

SSES_for <-forecast(SSES,h=35, interval="prediction", xreg = methane_test_ts[,5])

forecast_SSES_accuracy <- accuracy(SSES$forecast, methane_test_ts[,1])

plot(SSES_for)

print(forecast_SSES_accuracy)

#write to csv
write.csv(SSES_for$mean,
          "./Forecasts/Methane/ssesExoSE.csv",
          row.names = FALSE)

```


# Combination of results

```{r echo=FALSE}

forecast_arima$mean

forecast_arima_fourier$mean

forecast_sarima$mean

tbats_forecast$mean

NN_for$mean

stl_arima_forecast$mean

stl_arima_forecast

SSES$forecast

```



# A template for looping

```{r echo=FALSE}

variables <- list(c(2),c(2,3),c(2,4),c(2,5),c(2,3,4),c(2,3,5),c(2,4,5),c(2,3,4,5),c(3),c(3,4),c(3,5),c(3,4,5),c(4),c(4,5),c(5))

variables <- list(c(2),c(3),c(4),c(5))

result <- list()

for(variable in variables){

  #####
  # Your Code
  #####
  
  
  #####
  # Find the best variable
  #####
  
  result_list <- data.frame(Variable_colnum = variable, RMSE = xxxx_accuracy[,"MAPE"])
  
  result <- append(result,list(result_list))
  
  cat(variable,":\n")
  
  print(xxxxx_accuracy)

}

rmse_values <- sapply(result, function(x) x$RMSE)

best_index <- which.min(rmse_values)

best_result <- result[[best_index]]

print("The best combination is: \n")
print(best_result)

```










