---
title: "Forecasting models (Methane & Methane plus beef"
output: html_notebook
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(lubridate)
library(kableExtra)
library(smooth)
library(knitr)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 12), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## Deseason Data

```{r echo=FALSE}
#decompose the training series
decomposed <- decompose(methane_train_ts, type = "additive")
plot(decomposed)
```

```{r echo=FALSE}
#detrend and deseason the series
methane_trend <- decomposed$trend
methane_seasonality <- decomposed$seasonal
methane_detrend <- methane_train_ts - methane_trend
methane_deseasoned <- methane_train_ts - methane_seasonality
```

# Forecasting Models

## ARIMA

```{r echo=FALSE}
#autofit the sarima model
arima_autofit <- auto.arima(methane_deseasoned, 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0)
print(arima_autofit)

#get residuals
checkresiduals(arima_autofit)

#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = 36)

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima$mean, methane_test_ts)

#plot the forecast
plot(forecast_arima)
```

The ARIMA auto-fit was: ARIMA(5,2,4).

The residuals look pretty good. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## SARIMA

```{r echo=FALSE}
#autofit the sarima model
sarima_autofit <- auto.arima(methane_train_ts)
print(sarima_autofit)

#get residuals
checkresiduals(sarima_autofit)

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = 36)

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts)

#plot the forecast
plot(forecast_sarima)
```

The SARIMA auto-fit was: ARIMA(2,1,2)(1,1,1)[12].

The residuals look pretty good. There is one notable outlier. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## ARIMA with Fourier

```{r echo=FALSE}
#autofit the arima model
arima_fourier_autofit <- auto.arima(methane_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(methane_train_msts,
                                                 K=c(2,6)))
print(arima_fourier_autofit)

#get residuals
checkresiduals(arima_fourier_autofit)

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(methane_train_msts,
                                                K=c(2,6),
                                                h=36),
                                   h=365)
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts)

#plot the forecast
plot(forecast_arima_fourier)
```

## STL+ETS 

```{r}
#fit STL with ETS
stl_ets_fit <- stlm(methane_train_ts, s.window = "periodic", method = "ets")
print(stl_ets_fit$model)

#forecasting test data
stl_ets_forecast <- forecast(stl_ets_fit, h = length(methane_test_ts))

#accuracy assessment
stl_ets_accuracy <- accuracy(stl_ets_forecast$mean, methane_test_ts)

#visualization
plot(stl_ets_forecast)
```

## TBATS 

```{r}
#fit TBATS
tbats_fit <- tbats(methane_train_msts)
print(tbats_fit)
checkresiduals(tbats_fit)

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_ts))

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_ts)

#visualization
plot(tbats_forecast)
```

## Neuron Network

```{r}

NN_fit <- nnetar(methane_train_ts, p = 1, P = 7)

NN_for <- forecast(NN_fit, h=36)

forecast_NN_accuracy <- accuracy(NN_for$mean, methane_test_ts)

print(forecast_NN_accuracy)

plot(NN_for)

```

```{r}
NN_fourier_fit <- nnetar(methane_train_msts, p = 0, P = 7,
                 xreg=fourier(methane_train_msts,K=c(2,6)))

NN_fourier_for <- forecast(NN_fourier_fit, h=36, xreg=fourier(methane_train_msts,K=c(2,6), h=36))

forecast_NN_fourier_accuracy <- accuracy(NN_fourier_for$mean, methane_test_msts)

print(forecast_NN_fourier_accuracy)

plot(NN_fourier_for)

```

## State Space - Smooth

```{r}
SSES <- es(methane_train_ts,model="ZZZ",h=36,holdout=FALSE)

SSES_for <-forecast(SSES,h=36, interval="prediction")

plot(SSES_for)

forecast_SSES_accuracy <- accuracy(SSES$forecast, methane_test_ts)

print(forecast_SSES_accuracy)
```

## State Space - BSM

```{r}
SSBSM <- StructTS(methane_train_ts,
                    type="BSM",fixed=c(NA,NA,NA,NA))

SSBSM_for <- forecast(SSBSM,h=36)

plot(SSBSM_for)

forecast_SSBSM_accuracy <- accuracy(SSBSM_for$mean,methane_test_ts)

print(forecast_SSBSM_accuracy)

```

# Visualization
```{r}
methane_all <- read.csv("../Data/Processed/methane_all.csv")
methane_all_ts <- ts(
  methane_all$average,
  start = c(1983, 7),
  frequency = 12
)

arima_forecast_seasoned <- ts(
  forecast_arima$mean + decomposed$seasonal[length(methane_train_ts) + 1:36],
  start = c(2022, 1),
  frequency = 12
)


autoplot(methane_all_ts) +
  autolayer(arima_forecast_seasoned,
            series = "ARIMA (reseasoned)",
            PI = FALSE) +
  autolayer(forecast_sarima$mean,
            series = "SARIMA",
            PI = FALSE) +
  autolayer(forecast_arima_fourier$mean,
            series = "ARIMA + Fourier",
            PI = FALSE) +
  autolayer(stl_ets_forecast$mean,
            series = "STL + ETS",
            PI = FALSE) +
  autolayer(tbats_forecast$mean,
            series = "TBATS",
            PI = FALSE) +
  autolayer(NN_for$mean,
            series = "Neuron Network",
            PI = FALSE) +
  autolayer(NN_fourier_for$mean,
            series = "Neuron Network + Fourier",
            PI = FALSE) +
  autolayer(SSES_for$mean,
            series = "State Space - Smooth",
            PI = FALSE) +
  autolayer(SSBSM_for$mean,
            series = "State Space - BSM",
            PI = FALSE) +
  autolayer(window(methane_all_ts, start = c(2022, 1)),
            series = "True Data") +
  labs(title = "Methane Forecasts",
       y = "Methane Concentration (ppb)",
       x = "Time",
       col = "Models") +
  coord_cartesian(xlim = c(2021, 2024), ylim = c(1885, 1935))

```

```{r}
forecast_performance <- data.frame(rbind(
  forecast_arima_accuracy,
  forecast_sarima_accuracy,
  forecast_arima_fourier_accuracy,
  stl_ets_accuracy,
  tbats_accuracy,
  forecast_NN_accuracy,
  forecast_NN_fourier_accuracy,
  forecast_SSES_accuracy,
  forecast_SSBSM_accuracy
))

rownames(forecast_performance) <- c(
  "ARIMA (reseasoned)",
  "SARIMA",
  "ARIMA + Fourier",
  "STL + ETS",
  "TBATS",
  "Neuron Network",
  "Neuron Network + Fourier",
  "State Space - Smooth",
  "State Space - BSM"
)


kable(forecast_performance,
      caption = "Forecast Accuracy Metrics for All Models",
      digits = 3,
      booktabs = TRUE)
```

# ARIMAX model with beef predictor

```{r}
#import data
beef_production <- read.csv("../Data/Raw/beef_buffalo_production.csv", skip = 23,
                            header = FALSE)

#set column names
colnames(beef_production) <- c("Entity", "Code", "Year", "Beef_Production(Tons)")

write.csv(beef_production, 
          file = "../Data/Processed/beef_production.csv",
          row.names=FALSE)
```


```{r}
#create timeseries object
beef_ts <-ts(beef_production$`Beef_Production(Tons)`, 
                start = 1983, 
                end = 2023, 
                frequency = 1)

#plot timeseries
autoplot(beef_ts)+
  labs(title = "Global Beef Production from 1961 to 2023",
       y = "Beef (tons)")
```

```{r}
methane <- read.csv("../Data/Processed/methane_all.csv")
methane_yearly <- methane %>% 
  filter(year < 2024) %>% 
  group_by(year) %>% 
  summarize(yearly_average = mean(average))
```

```{r}
methane_beef <- cbind(methane_yearly, beef_production) 
methane_beef <- methane_beef %>%
  select(year, yearly_average, `Beef_Production(Tons)`)

```

```{r}
cor(methane_beef$`Beef_Production(Tons)`, methane_beef$yearly_average)

ggplot(methane_beef, aes(`Beef_Production(Tons)`, yearly_average)) + 
  geom_point()
```
```{r}
methane_beef_train <- subset(methane_beef, year <= 2021)
methane_beef_test <- subset(methane_beef, year > 2021)

beef_train <- ts(methane_beef_train$`Beef_Production(Tons)`, start = 1983)
methane_train <- ts(methane_beef_train$yearly_average, start = 1983)

beef_test <- ts(methane_beef_test$`Beef_Production(Tons)`, start = 2022)
methane_test <- ts(methane_beef_test$yearly_average, start = 2022)
```


```{r}
arimax_fit <- auto.arima(methane_train, xreg = beef_train)
summary(arimax_fit)
checkresiduals(arimax_fit)
```

```{r}
forecast_arimax <- forecast(arimax_fit, xreg = beef_test, h = length(methane_test))

autoplot(forecast_arimax) +
  autolayer(methane_test, series = "Actual", PI = FALSE) +
  labs(title = "ARIMAX Forecast: Methane Emissions using Beef Production",
       y = "Methane Emissions",
       x = "Year") +
  guides(colour = guide_legend(title = "Legend"))

```

```{r}
accuracy(forecast_arimax$mean, methane_test)
```

```{r}
social_economic <- read.csv("../Data/Processed/social_economic_factors_monthly.csv")
social_train <- social_economic[1:473,]
social_test <- social_economic[474:497,]

methane_social_train <- ts(social_train$methane, frequency=12)
cpi_train <- ts(social_train$cpi, frequency=12)
export_train <- ts(social_train$export, frequency=12)
import_train <- ts(social_train$import, frequency=12)
ip_train <- ts(social_train$ip, frequency=12)

methane_social_test <- ts(social_test$methane, frequency=12)
cpi_test <- ts(social_test$cpi, frequency=12)
export_test <- ts(social_test$export, frequency=12)
import_test <- ts(social_test$import, frequency=12)
ip_test <- ts(social_test$ip, frequency=12)

```

```{r}
arimax_fit_cpi <- auto.arima(methane_social_train, xreg = cpi_train)
summary(arimax_fit_cpi)
checkresiduals(arimax_fit_cpi)
```

```{r}
forecast_arimax_cpi <- forecast(arimax_fit_cpi, xreg = cpi_test, h = length(methane_social_test))
```

```{r}
arimax_fit_export <- auto.arima(methane_social_train, xreg = export_train)
summary(arimax_fit_export)
checkresiduals(arimax_fit_export)
forecast_arimax_export <- forecast(arimax_fit_export, xreg = export_test, h = length(methane_social_test))

```
```{r}
arimax_fit_import <- auto.arima(methane_social_train, xreg = import_train)
summary(arimax_fit_import)
checkresiduals(arimax_fit_import)
forecast_arimax_import <- forecast(arimax_fit_import, xreg = import_test, h = length(methane_social_test))
```

```{r}
arimax_fit_ip <- auto.arima(methane_social_train, xreg = ip_train)
summary(arimax_fit_ip)
checkresiduals(arimax_fit_ip)
forecast_arimax_ip <- forecast(arimax_fit_ip, xreg = ip_test, h = length(methane_social_test))
```


```{r}
# Reset test and forecasts to start from time = 1
n_test <- length(social_test$methane)

# Actual test data
methane_social_test <- ts(social_test$methane, start = 1, frequency = 12)

# Forecasts aligned with test
forecast_cpi_ts <- ts(forecast_arimax_cpi$mean, start = 1, frequency = 12)
forecast_export_ts <- ts(forecast_arimax_export$mean, start = 1, frequency = 12)
forecast_import_ts <- ts(forecast_arimax_import$mean, start = 1, frequency = 12)
forecast_ip_ts <- ts(forecast_arimax_ip$mean, start = 1, frequency = 12)

autoplot(methane_social_test) +
  autolayer(forecast_ip_ts,
            series = "IP",
            PI = FALSE) +

  autolayer(window(methane_social_test),
            series = "True Data") +
  labs(title = "Methane Forecasts",
       y = "Methane Concentration (ppb)",
       x = "Time",
       col = "Models")

```


```{r}

# Accuracy for each model
acc_cpi <- accuracy(forecast_cpi_ts, methane_social_test)
acc_export <- accuracy(forecast_export_ts, methane_social_test)
acc_import <- accuracy(forecast_import_ts, methane_social_test)
acc_ip <- accuracy(forecast_ip_ts, methane_social_test)

# Combine into a single data frame
acc_all <- rbind(
  "ARIMAX CPI" = acc_cpi,
  "ARIMAX Export" = acc_export,
  "ARIMAX Import" = acc_import,
  "ARIMAX IP" = acc_ip
)

rownames(acc_all) <- c(
  "CPI ARIMAX",
  "Export ARIMAX",
  "Import ARIMAX",
  "IP ARIMAX"
)
# Display using kable
kable(acc_all, digits = 2, caption = "Accuracy of ARIMAX Forecasts on Test Set")

```