---
output: html_document
editor_options: 
  chunk_output_type: console
---
```{r include=FALSE}
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)
library(tidyverse)
library(tibble)
```

# Temperature
## Upload Previous Models
```{r}
#upload files
filenames <- list.files("./Forecasts/Temperature", full.names = TRUE)

# Extract base names without extensions (to use as object names)
object_names <- tools::file_path_sans_ext(basename(filenames))

# Read the CSVs into a named list
forecasts <- setNames(lapply(filenames, read_csv), object_names)

#create df
forecast_df <- as.data.frame(lapply(forecasts, function(df) df[[1]]))

#Create objects in global environment.
list2env(forecasts, .GlobalEnv)

rm(filenames)

#upload test data
tempanom_test <- read.csv("./Data/Processed/tempanom_test.csv")
```

## Averaging
Averaging every combination of the models will result in 

```{r}
# Get column names
col_names <- names(forecast_df)

# Initialize empty list to hold averaged columns
avg_forecasts <- list()

# Loop over sizes of combinations (from 2 to 9 models)
for (r in 1:length(col_names)) {
  combos <- combn(col_names, r, simplify = FALSE)
  
  for (combo in combos) {
    combo_name <- paste(combo, collapse = "_")
    avg_forecasts[[combo_name]] <- rowMeans(forecast_df[, combo, drop = FALSE])
  }
}

# Convert to a data frame
avg_forecasts_df <- as.data.frame(avg_forecasts)
```

Get the performance of each combination and save the top 10 to a separate dataframe

```{r}
avg_forecasts_accuracy <- data.frame(matrix(ncol = 5, nrow = 0))

#provide column names
colnames(avg_forecasts_accuracy) <- c('ME', 'RSME', 'MAE', "MPE", "MAPE")

#calculate accuracy statistics
for (i in 1:length(avg_forecasts_df)) {
  avg_forecasts_accuracy[i,] <- accuracy(avg_forecasts_df[,i], tempanom_test$temp_anomaly)
}

#set row names
row.names(avg_forecasts_accuracy) <- colnames(avg_forecasts_df)
```

## Extracting Best Forecasts and save

```{r}
#get best averaged forecasts
avg_forecasts_accuracy_best <- rbind(avg_forecasts_accuracy) %>%
  arrange(abs(MAPE)) %>%
  head(5) #get 5 best forecasts

forecast_save <- rownames_to_column(avg_forecasts_accuracy_best, var = "Model")

#write into csv
write.csv(forecast_save, 
         file = here("Forecasts/temperature_accuracy.csv"),
         row.names = TRUE)

#save averaged forecasts
#grab names of all averaged forecasts
avg_names <- forecast_save$Model

#export averaged forecasts
for (i in seq_along(avg_names)) {
  filepath <- paste0("Forecasts/Temperature_Averages/", avg_names[i], ".csv")
  
  df <- avg_forecasts_df[[avg_names[i]]]
  
  write.csv(df, 
          file = here(filepath),
          row.names = FALSE)
}

```

# Methane
## Upload Previous Models
```{r}
#clear environment from previous
rm(list=ls())

#upload files
filenames <- list.files("./Forecasts/Methane", full.names = TRUE)

# Extract base names without extensions (to use as object names)
object_names <- tools::file_path_sans_ext(basename(filenames))

# Read the CSVs into a named list
forecasts <- setNames(lapply(filenames, read_csv), object_names)

#create df
forecast_df <- as.data.frame(lapply(forecasts, function(df) df[[1]]))

#Create objects in global environment.
list2env(forecasts, .GlobalEnv)

rm(filenames)

#upload test data
methane_test <- read.csv("./Data/Processed/methane_test.csv")
```

## Averaging
Averaging every combination of the models will result in 

```{r}
# Get column names
col_names <- names(forecast_df)

# Initialize empty list to hold averaged columns
avg_forecasts <- list()

# Loop over sizes of combinations (from 2 to 9 models)
for (r in 1:length(col_names)) {
  combos <- combn(col_names, r, simplify = FALSE)
  
  for (combo in combos) {
    combo_name <- paste(combo, collapse = "_")
    avg_forecasts[[combo_name]] <- rowMeans(forecast_df[, combo, drop = FALSE])
  }
}

# Convert to a data frame
avg_forecasts_df <- as.data.frame(avg_forecasts)
```

Get the performance of each combination and save the top 10 to a separate dataframe

```{r}
avg_forecasts_accuracy <- data.frame(matrix(ncol = 5, nrow = 0))

#provide column names
colnames(avg_forecasts_accuracy) <- c('ME', 'RSME', 'MAE', "MPE", "MAPE")

#calculate accuracy statistics
for (i in 1:length(avg_forecasts_df)) {
  avg_forecasts_accuracy[i,] <- accuracy(avg_forecasts_df[,i], methane_test$average)
}

#set row names
row.names(avg_forecasts_accuracy) <- colnames(avg_forecasts_df)
```

## Extracting Best Forecasts and save

```{r}
#get best averaged forecasts
avg_forecasts_accuracy_best <- rbind(avg_forecasts_accuracy) %>%
  arrange(abs(MAPE)) %>%
  head(5) #get 5 best forecasts

forecast_save <- rownames_to_column(avg_forecasts_accuracy_best, var = "Model")

#write into csv
write.csv(forecast_save, 
         file = here("Forecasts/methane_accuracy.csv"),
         row.names = TRUE)

#save averaged forecasts
#grab names of all averaged forecasts
avg_names <- forecast_save$Model

#export averaged forecasts
for (i in seq_along(avg_names)) {
  filepath <- paste0("Forecasts/Methane_Averages/", avg_names[i], ".csv")
  
  df <- avg_forecasts_df[[avg_names[i]]]
  
  write.csv(df, 
          file = here(filepath),
          row.names = FALSE)
}

```
