---
title: "R Notebook"
output: html_notebook
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(lubridate)
library(kableExtra)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 12), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## Deseason Data

```{r echo=FALSE}
#decompose the training series
decomposed <- decompose(methane_train_ts, type = "additive")
plot(decomposed)
```

```{r echo=FALSE}
#detrend and deseason the series
methane_trend <- decomposed$trend
methane_seasonality <- decomposed$seasonal
methane_detrend <- methane_train_ts - methane_trend
methane_deseasoned <- methane_train_ts - methane_seasonality
```

# Forecasting Models

## ARIMA

```{r echo=FALSE}
#autofit the sarima model
arima_autofit <- auto.arima(methane_deseasoned, 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0)
print(arima_autofit)

#get residuals
checkresiduals(arima_autofit)

#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = 36)

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima$mean, methane_test_ts)

#plot the forecast
plot(forecast_arima)
```

The ARIMA auto-fit was: ARIMA(5,2,4).

The residuals look pretty good. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## SARIMA

```{r echo=FALSE}
#autofit the sarima model
sarima_autofit <- auto.arima(methane_train_ts)
print(sarima_autofit)

#get residuals
checkresiduals(sarima_autofit)

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = 36)

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts)

#plot the forecast
plot(forecast_sarima)
```

The SARIMA auto-fit was: ARIMA(2,1,2)(1,1,1)[12].

The residuals look pretty good. There is one notable outlier. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## ARIMA with Fourier

```{r echo=FALSE}
#autofit the arima model
arima_fourier_autofit <- auto.arima(methane_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(methane_train_msts,
                                                 K=c(2,6)))
print(arima_fourier_autofit)

#get residuals
checkresiduals(arima_fourier_autofit)

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(methane_train_msts,
                                                K=c(2,6),
                                                h=36),
                                   h=365)
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts)

#plot the forecast
plot(forecast_arima_fourier)
```

## STL+ETS 

```{r}
#fit STL with ETS
stl_ets_fit <- stlm(methane_train_ts, s.window = "periodic", method = "ets")
print(stl_ets_fit$model)

#forecasting test data
stl_ets_forecast <- forecast(stl_ets_fit, h = length(methane_test_ts))

#accuracy assessment
stl_ets_accuracy <- accuracy(stl_ets_forecast$mean, methane_test_ts)

#visualization
plot(stl_ets_forecast)
```

## TBATS 

```{r}
#fit TBATS
tbats_fit <- tbats(methane_train_msts)
print(tbats_fit)
checkresiduals(tbats_fit)

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_ts))

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_ts)

#visualization
plot(tbats_forecast)
```


# Visualization
```{r}
methane_all <- read.csv("./Data/Processed/methane_all.csv")
methane_all_ts <- ts(
  methane_all$average,
  start = c(1983, 7),
  frequency = 12
)

arima_forecast_seasoned <- ts(
  forecast_arima$mean + decomposed$seasonal[length(methane_train_ts) + 1:36],
  start = c(2022, 1),
  frequency = 12
)


autoplot(methane_all_ts) +
  autolayer(arima_forecast_seasoned,
            series = "ARIMA (reseasoned)",
            PI = FALSE) +
  autolayer(forecast_sarima$mean,
            series = "SARIMA",
            PI = FALSE) +
  autolayer(forecast_arima_fourier$mean,
            series = "ARIMA + Fourier",
            PI = FALSE) +
  autolayer(stl_ets_forecast$mean,
            series = "STL + ETS",
            PI = FALSE) +
  autolayer(tbats_forecast$mean,
            series = "TBATS",
            PI = FALSE) +
  autolayer(window(methane_full_ts, start = c(2022, 1)),
            series = "True Data") +
  labs(title = "Methane Forecasts",
       y = "Methane Concentration (ppb)",
       x = "Time",
       col = "Models") +
  coord_cartesian(xlim = c(2021, 2024), ylim = c(1885, 1935))

```

```{r}
forecast_performance <- data.frame(rbind(
  forecast_arima_accuracy,
  forecast_sarima_accuracy,
  forecast_arima_fourier_accuracy,
  stl_ets_accuracy,
  tbats_accuracy
))

rownames(forecast_performance) <- c(
  "ARIMA (reseasoned)",
  "SARIMA",
  "ARIMA + Fourier",
  "STL + ETS",
  "TBATS"
)


kable(forecast_performance,
      caption = "Forecast Accuracy Metrics for All Models",
      digits = 3,
      booktabs = TRUE)
```

