---
title: "Methane Forecasting"
output: 
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 12), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## Deseason Data

```{r echo=FALSE}
#decompose the training series
decomposed <- decompose(methane_train_ts, type = "additive")
plot(decomposed)
```

```{r echo=FALSE}
#detrend and deseason the series
methane_trend <- decomposed$trend
methane_seasonality <- decomposed$seasonal
methane_detrend <- methane_train_ts - methane_trend
methane_deseasoned <- methane_train_ts - methane_seasonality

detrend_plot <- autoplot(methane_detrend, ylim=c(-30,30), 
                         main="Detrended Plot") + ylab("Methane (ppb)")
deseasoned_plot <- autoplot(methane_deseasoned,
                            main="Deseasoned Plot") + ylab("Methane (ppb)")
plot_grid(detrend_plot, deseasoned_plot)
```

## ACF and PACF

```{r echo=FALSE, warning=FALSE}
#check acf and pacf
ACF_Plot <- autoplot(Acf(methane_train_ts, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="ACF plot")
PACF_Plot <- autoplot(Pacf(methane_train_ts, lag=40, plot = FALSE), 
                      ylim=c(-0.5,1), 
                      main="PACF plot")
plot_grid(ACF_Plot, PACF_Plot)
```

```{r echo=FALSE}
#check ACF and PACF of detrended and deseasoned series
ACF_detrend <- autoplot(Acf(methane_detrend, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Detrended ACF plot")

ACF_deseason <- autoplot(Acf(methane_deseasoned, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Deseasoned ACF plot")

PACF_detrend <- autoplot(Pacf(methane_detrend, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Detrended PACF plot")

PACF_deseason <- autoplot(Pacf(methane_deseasoned, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Deseasoned PACF plot")

plot_row1 <- plot_grid(ACF_Plot, ACF_detrend, ACF_deseason, nrow=1, ncol=3)
plot_row2 <- plot_grid(PACF_Plot, PACF_detrend, PACF_deseason, nrow=1, ncol=3)
plot_grid(plot_row1, plot_row2, nrow=2, ncol=1)
```

## Stationarity Tests

```{r echo=FALSE}
#Mann-Kendall test
print("Results from Seasonal Mann-Kendall Test for orginal data")
summary(SeasonalMannKendall(methane_train_ts))

print("Results from Mann-Kendall Test for deseasoned data")
summary(MannKendall(methane_deseasoned))
```

```{r echo=FALSE}
#ADF test
print("Results from ADF test for original data")
print(adf.test(methane_train_ts,alternative = "stationary"))
```

```{r echo=FALSE}
#Spearman Correlation
print("Results from Spearman Correlation")
print(cor.test(methane_train_ts,c(1:462),method="spearman"))
```

# Forecasting Models

## ARIMA

```{r echo=FALSE}
#autofit the sarima model
arima_autofit <- auto.arima(methane_deseasoned, 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0)
print(arima_autofit)

#get residuals
checkresiduals(arima_autofit)

#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = 36)

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima$mean, methane_test_ts)

#plot the forecast
plot(forecast_arima)
```

The ARIMA auto-fit was: ARIMA(5,2,4).

The residuals look pretty good. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## SARIMA

```{r echo=FALSE}
#autofit the sarima model
sarima_autofit <- auto.arima(methane_train_ts)
print(sarima_autofit)

#get residuals
checkresiduals(sarima_autofit)

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = 36)

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts)

#plot the forecast
plot(forecast_sarima)
```

The SARIMA auto-fit was: ARIMA(2,1,2)(1,1,1)[12].

The residuals look pretty good. There is one notable outlier. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## ARIMA with Fourier

```{r echo=FALSE}
#autofit the arima model
arima_fourier_autofit <- auto.arima(methane_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(methane_train_msts,
                                                 K=c(2,6)))
print(arima_fourier_autofit)

#get residuals
checkresiduals(arima_fourier_autofit)

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(methane_train_msts,
                                                K=c(2,6),
                                                h=36),
                                   h=365)
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts)

#plot the forecast
plot(forecast_arima_fourier)
```

# Performance Comparison

```{r echo=FALSE}
# Combine performance vectors as rows
forecast_performance <- data.frame(rbind(forecast_arima_accuracy, 
                                         forecast_sarima_accuracy,
                                         forecast_arima_fourier_accuracy))

# Set row names
rownames(forecast_performance) <- c("ARIMA",
                                    "SARIMA",
                                    "ARIMA with Fourier")
```

```{r echo=FALSE}
kable(forecast_performance,
      #format = "latex",  # Works for PDF output
      caption = "Forecast Accuracy",
      digits = 3,
      booktabs = TRUE)
```
