---
title: "Methane Forecasting"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")
methane_all <- read.csv("./Data/Processed/methane_all.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 11), 
                frequency = 12)

methane_all_ts <-ts(methane_all$average, 
                start = c(1983,7), 
                end = c(2024, 11), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## Deseason Data

```{r echo=FALSE}
#decompose the training series
decomposed <- decompose(methane_train_ts, type = "additive")
plot(decomposed)
```

```{r echo=FALSE}
#detrend and deseason the series
methane_trend <- decomposed$trend
methane_seasonality <- decomposed$seasonal
methane_detrend <- methane_train_ts - methane_trend
methane_deseasoned <- methane_train_ts - methane_seasonality

detrend_plot <- autoplot(methane_detrend, ylim=c(-30,30), 
                         main="Detrended Plot") + ylab("Methane (ppb)")
deseasoned_plot <- autoplot(methane_deseasoned,
                            main="Deseasoned Plot") + ylab("Methane (ppb)")
plot_grid(detrend_plot, deseasoned_plot)
```

## ACF and PACF

```{r echo=FALSE, warning=FALSE}
#check acf and pacf
ACF_Plot <- autoplot(Acf(methane_train_ts, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="ACF plot")
PACF_Plot <- autoplot(Pacf(methane_train_ts, lag=40, plot = FALSE), 
                      ylim=c(-0.5,1), 
                      main="PACF plot")
plot_grid(ACF_Plot, PACF_Plot)
```

```{r echo=FALSE}
#check ACF and PACF of detrended and deseasoned series
ACF_detrend <- autoplot(Acf(methane_detrend, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Detrended ACF plot")

ACF_deseason <- autoplot(Acf(methane_deseasoned, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Deseasoned ACF plot")

PACF_detrend <- autoplot(Pacf(methane_detrend, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Detrended PACF plot")

PACF_deseason <- autoplot(Pacf(methane_deseasoned, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="Deseasoned PACF plot")

plot_row1 <- plot_grid(ACF_Plot, ACF_detrend, ACF_deseason, nrow=1, ncol=3)
plot_row2 <- plot_grid(PACF_Plot, PACF_detrend, PACF_deseason, nrow=1, ncol=3)
plot_grid(plot_row1, plot_row2, nrow=2, ncol=1)
```

## Stationarity Tests

```{r echo=FALSE}
#Mann-Kendall test
print("Results from Seasonal Mann-Kendall Test for orginal data")
summary(SeasonalMannKendall(methane_train_ts))

print("Results from Mann-Kendall Test for deseasoned data")
summary(MannKendall(methane_deseasoned))
```

```{r echo=FALSE}
#ADF test
print("Results from ADF test for original data")
print(adf.test(methane_train_ts,alternative = "stationary"))
```

```{r echo=FALSE}
#Spearman Correlation
print("Results from Spearman Correlation")
print(cor.test(methane_train_ts,c(1:462),method="spearman"))
```

# Forecasting Models

## ARIMA

```{r echo=FALSE}
#autofit the sarima model
arima_autofit <- auto.arima(methane_deseasoned, 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0)
print(arima_autofit)

#get residuals
checkresiduals(arima_autofit)

#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = length(methane_test_ts)) 

#add seasonality
forecast_arima_seasoned <- ts(
  forecast_arima$mean + decomposed$seasonal[length(methane_train_ts) + 1:36],
  start = c(2022, 1),
  frequency = 12)

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima_seasoned[1:35], methane_test_ts)

#plot the forecast
#autoplot(forecast_arima_seasoned)
```

The ARIMA auto-fit was: ARIMA(5,2,4).

The residuals look pretty good. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## SARIMA

```{r echo=FALSE}
#autofit the sarima model
sarima_autofit <- auto.arima(methane_train_ts)
print(sarima_autofit)

#get residuals
checkresiduals(sarima_autofit)

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = length(methane_test_ts))

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, methane_test_ts)

#plot the forecast
plot(forecast_sarima)
```

The SARIMA auto-fit was: ARIMA(2,1,2)(1,1,1)[12].

The residuals look pretty good. There is one notable outlier. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

## ARIMA with Fourier

```{r echo=FALSE}
#autofit the arima model
arima_fourier_autofit <- auto.arima(methane_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(methane_train_msts,
                                                 K=c(2,6)))
print(arima_fourier_autofit)

#get residuals
checkresiduals(arima_fourier_autofit)

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(methane_train_msts,
                                                K=c(2,6),
                                                h=length(methane_test_ts)),
                                   h=length(methane_test_ts))
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, methane_test_msts)

#plot the forecast
plot(forecast_arima_fourier)
```

##STL+ETS

```{r}
#fit STL with ETS
stl_ets_fit <- stlm(methane_train_ts, s.window = "periodic", method = "ets")
print(stl_ets_fit$model)

#forecasting test data
stl_ets_forecast <- forecast(stl_ets_fit, h = length(methane_test_ts))

#accuracy assessment
stl_ets_accuracy <- accuracy(stl_ets_forecast$mean, methane_test_ts)

#visualization
plot(stl_ets_forecast)
```

##TBATS

```{r}
#fit TBATS
tbats_fit <- tbats(methane_train_msts)
print(tbats_fit)
checkresiduals(tbats_fit)

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_ts))

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_ts)

#visualization
plot(tbats_forecast)
```

## Neuron Network

```{r}

NN_fit <- nnetar(methane_train_ts, p = 1, P = 7)

NN_for <- forecast(NN_fit, h = length(methane_test_ts))

forecast_NN_accuracy <- accuracy(NN_for$mean, methane_test_ts)

print(forecast_NN_accuracy)

plot(NN_for)

```

```{r}

NN_fourier_fit <- nnetar(methane_train_msts, p = 0, P = 7,
                 xreg=fourier(methane_train_msts,K=c(2,6)))

NN_fourier_for <- forecast(NN_fourier_fit, h=length(methane_test_ts), xreg=fourier(methane_train_msts,K=c(2,6), h=length(methane_test_ts)))

forecast_NN_fourier_accuracy <- accuracy(NN_fourier_for$mean, methane_test_msts)

print(forecast_NN_fourier_accuracy)

plot(NN_fourier_for)

```

## State Space - Smooth

```{r}

SSES <- es(methane_train_ts,
           model="ZZZ",
           h=length(methane_test_ts),
           holdout=FALSE)

SSES_for <-forecast(SSES,
                    h = length(methane_test_ts), 
                    interval="prediction")

plot(SSES_for)

forecast_SSES_accuracy <- accuracy(SSES$forecast, methane_test_ts)

print(forecast_SSES_accuracy)

```

## State Space - BSM

```{r}

SSBSM <- StructTS(methane_train_ts,
                    type="BSM",fixed=c(NA,NA,NA,NA))

SSBSM_for <- forecast(SSBSM,h=length(methane_test_ts))

plot(SSBSM_for)

forecast_SSBSM_accuracy <- accuracy(SSBSM_for$mean,methane_test_ts)

print(forecast_SSBSM_accuracy)

```


# Performance Comparison

```{r echo=FALSE}
# Combine performance vectors as rows
forecast_performance <- data.frame(rbind(forecast_arima_accuracy, 
                                         forecast_sarima_accuracy,
                                         forecast_arima_fourier_accuracy,
                                         stl_ets_accuracy,
                                         tbats_accuracy,
                                         forecast_NN_accuracy, 
                                         forecast_NN_fourier_accuracy,
                                         forecast_SSES_accuracy,
                                         forecast_SSBSM_accuracy))

# Set row names
rownames(forecast_performance) <- c("ARIMA",
                                    "SARIMA",
                                    "ARIMA w/ Fourier",
                                    "STL + ETS",
                                    "TBATS",
                                    "Neural Network",
                                    "Neural Network w/ fourier",
                                    "State Space w/ Exponential smoothing",
                                    "State Space w/ BSM")
```

```{r echo=FALSE}
kable(forecast_performance[,1:6], 
      caption = "Forecast Accuracy",
      digits = 3)
```

# Averaging
## Averaged Models

Averaging every combination of the models will result in 

```{r}
#create a dataframe of all of the forecasts
forecasts <- data.frame(
  "arima" = forecast_arima$mean,
  "sarima" = forecast_sarima$mean,
  "arimaFourier" = forecast_arima_fourier$mean,
  "stlEts" = stl_ets_forecast$mean,
  "tbats" = tbats_forecast$mean,
  "NN" = NN_for$mean,
  "NNFourier" = NN_fourier_for$mean,
  "SSSmooth" = SSES_for$mean,
  "SSBSM" = SSBSM_for$mean)

# Get column names
col_names <- names(forecasts)

# Initialize empty list to hold averaged columns
avg_forecasts <- list()

# Loop over sizes of combinations (from 2 to 9 models)
for (r in 2:length(col_names)) {
  combos <- combn(col_names, r, simplify = FALSE)
  
  for (combo in combos) {
    combo_name <- paste(combo, collapse = "_")
    avg_forecasts[[combo_name]] <- rowMeans(forecasts[, combo, drop = FALSE])
  }
}

# Convert to a data frame
avg_forecasts_df <- as.data.frame(avg_forecasts)
```

Get the performance of each combination and save the top 10 to a separate dataframe

```{r}
avg_forecasts_accuracy <- data.frame(matrix(ncol = 7, nrow = 0))

#provide column names
colnames(avg_forecasts_accuracy) <- c('ME', 'RSME', 'MAE', "MPE", "MAPE", "ACF1", "Theil.s.U")

#calculate accuracy statistics
for (i in 1:length(avg_forecasts_df)) {
  avg_forecasts_accuracy[i,] <- accuracy(avg_forecasts_df[,i], methane_test_ts)
}

#set row names
row.names(avg_forecasts_accuracy) <- colnames(avg_forecasts_df)

#get best averaged forecasts
avg_forecasts_accuracy_best <- avg_forecasts_accuracy %>%
  arrange(abs(MAPE)) %>%
  head(10)
```

# All Forecasts Plot

```{r}
autoplot(methane_all_ts) +
  autolayer(stl_ets_forecast, PI=FALSE, series="STL+ETS", alpha = 0.75) +
  autolayer(forecast_arima_seasoned, PI=FALSE, series="ARIMA", alpha = 0.75) +
  autolayer(forecast_sarima, PI=FALSE, series="SARIMA", alpha = 0.75) +
  autolayer(forecast_arima_fourier, PI=FALSE, series="ARIMA + Fourier", alpha = 0.75) +
  autolayer(tbats_forecast,PI=FALSE, series="TBATS", alpha = 0.75) +
  autolayer(NN_for,PI=FALSE, series="Neural Network", alpha = 0.75) +
  autolayer(NN_fourier_for,PI=FALSE, series="Neural Network w/ Fourier", alpha = 0.75) +
  autolayer(SSES_for,PI=FALSE, series="State Space w/ Exponential smoothing", alpha = 0.75) +
  autolayer(SSBSM_for,PI=FALSE, series="State Space w/ BSM", alpha = 0.75)+
  labs(x="Time",
       y="Methane Concentration (ppb)",
       title = "Methane Forecasts") +
  guides(colour=guide_legend(title="Forecast"))+
  xlim(c(2021 + 5/12, 2024 + 10/12))+
  ylim(c(1880,1960))
```

