---
title: "Temperature Forecasting"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)
library(randomForest)
library(xts)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data for methane
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")
methane_all <- read.csv("./Data/Processed/methane_all.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 11), 
                frequency = 12)

methane_all_ts <-ts(methane_all$average, 
                start = c(1983,7), 
                end = c(2024, 11), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))

#upload test and training data for temperature
tempanom_train <- read.csv("./Data/Processed/tempanom_training.csv")
tempanom_test <- read.csv("./Data/Processed/tempanom_test.csv")
tempanom_all <- read.csv("./Data/Processed/tempanom_all.csv")

#create ts object
tempanom_train_ts <-ts(tempanom_train$temp_anomaly, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

tempanom_test_ts <-ts(tempanom_test$temp_anomaly, 
                start = c(2022,1), 
                end = c(2024, 11), 
                frequency = 12)

tempanom_all_ts <-ts(c(tempanom_train_ts,tempanom_test_ts), 
                     start=start(tempanom_train_ts), 
                     frequency=frequency(tempanom_train_ts))

#create msts
tempanom_train_msts <- msts(tempanom_train$temp_anomaly, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

tempanom_test_msts <- msts(tempanom_test$temp_anomaly, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## ACF and PACF

```{r echo=FALSE, warning=FALSE}
#check acf and pacf
ACF_Plot <- autoplot(Acf(tempanom_train_ts, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="ACF plot")
PACF_Plot <- autoplot(Pacf(tempanom_train_ts, lag=40, plot = FALSE), 
                      ylim=c(-0.5,1), 
                      main="PACF plot")

#create plot grid
plot_grid(ACF_Plot, PACF_Plot)
```

# Correlations

```{r}
#create df
methane_temp <- data.frame(methane = methane_train$average,
                           tempanom = tempanom_train$temp_anomaly)

#determine correlation within the df
methane_temp_cor <- cor(methane_temp)
```

# Model Fitting
## Endogenous Only
### ------ ARIMA

```{r echo=FALSE}
#autofit the sarima model 
arima_autofit <- auto.arima(methane_deseasoned, 
                            max.D = 0, 
                            max.P = 0, 
                            max.Q = 0)
print(arima_autofit)

#get residuals
checkresiduals(arima_autofit)

#create the sarima forecast with the autofit
forecast_arima <- forecast(object = arima_autofit, h = length(methane_test_ts)) 

#add seasonality
forecast_arima <- forecast_arima$mean + rep(tail(methane_seasonality, 12), 
                                                length.out = length(methane_test_ts))

#calculate accuracy metrics
forecast_arima_accuracy <- accuracy(forecast_arima, methane_test_ts)

#plot the forecast
autoplot(forecast_arima)
```

The ARIMA auto-fit was: ARIMA(5,2,4).

The residuals look pretty good. Otherwise, the residuals are primarily not significant in the ACF plot and are normally distributed.

### SARIMA

```{r echo=FALSE}
#autofit the sarima model
sarima_autofit <- auto.arima(tempanom_train_ts)

#create the sarima forecast with the autofit
forecast_sarima <- forecast(object = sarima_autofit, h = length(tempanom_test_ts))

#calculate accuracy metrics
forecast_sarima_accuracy <- accuracy(forecast_sarima$mean, tempanom_test_ts)

#write to csv
write.csv(forecast_sarima$mean, 
         file = "./Forecasts/Temperature/sarima.csv",
         row.names = FALSE)
```

### ARIMA with Fourier

```{r echo=FALSE}
#autofit the arima model
arima_fourier_autofit <- auto.arima(tempanom_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(tempanom_train_msts,
                                                 K=c(2,4)))

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(tempanom_train_msts,
                                                K=c(2,4),
                                                h=length(tempanom_test_msts)),
                                   h=length(tempanom_test_msts))
  
#calculate accuracy metrics
forecast_arima_fourier_accuracy <- accuracy(forecast_arima_fourier$mean, tempanom_test_msts)

#write to csv
write.csv(forecast_arima_fourier$mean, 
         file = "./Forecasts/Temperature/arimaFourier.csv",
         row.names = FALSE)
```

### ------STL+ETS

```{r}
#fit STL with ETS
stl_ets_fit <- stlm(methane_train_ts, s.window = "periodic", method = "ets")
print(stl_ets_fit$model)

#forecasting test data
stl_ets_forecast <- forecast(stl_ets_fit, h = length(methane_test_ts))

#accuracy assessment
stl_ets_accuracy <- accuracy(stl_ets_forecast$mean, methane_test_ts)

#visualization
plot(stl_ets_forecast)
```

###-------------TBATS

```{r}
#fit TBATS
tbats_fit <- tbats(methane_train_msts)
print(tbats_fit)
checkresiduals(tbats_fit)

#forecasting test data
tbats_forecast <- forecast(tbats_fit, h = length(methane_test_ts))

#accuracy assessment
tbats_accuracy <- accuracy(tbats_forecast$mean, methane_test_ts)

#visualization
plot(tbats_forecast)
```

###------------ Neuron Network

```{r}

NN_fit <- nnetar(methane_train_ts, p = 1, P = 7)

NN_for <- forecast(NN_fit, h = length(methane_test_ts))

forecast_NN_accuracy <- accuracy(NN_for$mean, methane_test_ts)

print(forecast_NN_accuracy)

plot(NN_for)

```

## Exogenous Variables

### SARIMAX

```{r}
#autofit arimax
arimax_fit <- auto.arima(tempanom_train_ts,
                         xreg = methane_train_ts)

#create forecast
forecast_arimax <- forecast(arimax_fit, 
                            xreg = methane_test_ts, 
                            h = length(methane_test))

#write into csv
write.csv(forecast_arimax$mean, 
         file = here("Forecasts/Temperature/sarimax.csv"),
         row.names = FALSE)

#calculate accuracy metrics
forecast_arimax_accuracy <- accuracy(forecast_arimax$mean, tempanom_test_ts)

```

### ARIMAX with Fourier

```{r}
#autofit the arima model
arima_fourier_autofit <- auto.arima(tempanom_train_msts,
                                    seasonal=FALSE,
                                    lambda=0,
                                    xreg=fourier(methane_train_msts,
                                                 K=c(2,4)))

#create the arima forecast with the autofit
forecast_arima_fourier <- forecast(object = arima_fourier_autofit,
                                   xreg=fourier(methane_train_msts,
                                                K=c(2,4),
                                                h=length(tempanom_test_ts)),
                                   h=length(tempanom_test_ts))

#write into csv
write.csv(forecast_arima_fourier$mean, 
         file = here("Forecasts/Temperature/arimaxFourier.csv"),
         row.names = FALSE)
```



### Neural Network

```{r}
#autofit
nnetar_fit <- nnetar(tempanom_train_msts, 
              xreg = methane_train_msts,
              repeats = 10)

#create forecast
forecast_nnetar <- forecast(nnetar_fit, 
                            xreg = methane_test_msts, 
                            h = length(methane_test))

#write into csv
write.csv(forecast_nnetar$mean, 
         file = here("Forecasts/Temperature/nnetarExo.csv"),
         row.names = FALSE)

#calculate accuracy metrics
forecast_nnetar_accuracy <- accuracy(forecast_nnetar$mean, tempanom_test_msts)
```

### Random Forest

```{r}
#create dataframe for test
rf_train_df <- data.frame("tempanom" = tempanom_train$temp_anomaly,
                          "methane" = methane_train$average)

rf_test_df <- data.frame("methane" = methane_test$average)

# Fit a Random Forest model
rf_fit <- randomForest(tempanom ~ ., data = rf_train_df, ntree = 100)

# Make predictions on the test data
forecast_rf <- predict(rf_fit, newdata = rf_test_df)

#write into csv
write.csv(forecast_rf, 
         file = here("Forecasts/Temperature/rfExo.csv"),
         row.names = FALSE)

#calculate accuracy metrics
forecast_rf_accuracy <- accuracy(forecast_rf, tempanom_test_msts)
```

# Performance Comparison

```{r echo=FALSE}
# Combine performance vectors as rows
forecast_performance <- data.frame(rbind(forecast_arimax_accuracy, 
                                         forecast_nnetar_accuracy,
                                         forecast_rf_accuracy))

# Set row names
rownames(forecast_performance) <- c("ARIMAX",
                                    "NNETAR",
                                    "Random Forest")
```

```{r echo=FALSE}
kable(forecast_performance[,1:5], 
      caption = "Forecast Accuracy",
      digits = 3)
```

# Plot

```{r}
# Convert predictions to a ts object
rf_ts <- ts(forecast_rf, start = start(tempanom_test_ts), frequency = frequency(tempanom_test_ts))

# Plot properly
autoplot(tempanom_train_ts) +
  geom_line(data = tempanom_test_ts, colour = "grey40") +
  autolayer(forecast_nnetar, PI = FALSE, series = "NNETAR") +
  autolayer(forecast_arimax, PI = FALSE, series = "ARIMAX") +
  autolayer(rf_ts, series = "Random Forest", size = 1, color = "firebrick") +
  labs(x = "Time",
       y = "Temperature Anomaly (*C)",
       title = "Temperature Anomaly over Time") +
  guides(colour = guide_legend(title = "Forecast")) +
  xlim(c(2010, 2025)) +
  ylim(c(0.4, 1.5))


```

# Averaging
## Averaged Models

Averaging every combination of the models will result in 

```{r}
#create a dataframe of all of the forecasts
forecasts <- data.frame(
  "arimax" = forecast_arimax$mean,
  "NNETAR" = forecast_nnetar$mean,
  )

# Get column names
col_names <- names(forecasts)

# Initialize empty list to hold averaged columns
avg_forecasts <- list()

# Loop over sizes of combinations (from 2 to 9 models)
for (r in 2:length(col_names)) {
  combos <- combn(col_names, r, simplify = FALSE)
  
  for (combo in combos) {
    combo_name <- paste(combo, collapse = "_")
    avg_forecasts[[combo_name]] <- rowMeans(forecasts[, combo, drop = FALSE])
  }
}

# Convert to a data frame
avg_forecasts_df <- as.data.frame(avg_forecasts)
```


# Notes
Models to check out:

VAR (Vector AutoRegression) models
Random Forests, Gradient Boosting, and Neural Networks
https://stackoverflow.com/questions/51764339/how-to-predict-using-var-with-exogenous-variables-in-r 
LSTM

Seasonal Autoregressive Integrated Moving-Average with Exogenous Regressors (SARIMAX) with Exogenous Regressors (SARIMAX) is an extension of the SARIMA model that also includes the modeling of exogenous variables.

The SARIMAX method can also be used to model the subsumed models with exogenous variables, such as ARX, MAX, ARMAX, and ARIMAX.

The method is suitable for univariate time series with trend and/or seasonal components and exogenous variables.

Vector Autoregression Moving-Average with Exogenous Regressors (VARMAX) is an extension of the VARMA model that also includes the modeling of exogenous variables. It is a multivariate version of the ARMAX method.
