---
title: "Temperature Forecasting"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options:
  chunk_output_type: console
---

```{r chunks, include=FALSE}
#set up chunk options
knitr::opts_chunk$set(echo = TRUE,tidy.opts=list(width.cutoff=80), 
                      tidy=FALSE) 
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```

```{r libraries, include=FALSE}
#retrieve libraries
library(here)
library(ggplot2)
library(forecast)
library(cowplot)
library(Kendall)
library(tseries)
library(kableExtra)
library(dplyr)
library(smooth)

set.seed(797)
```

```{r theme, include=FALSE}
#create plot theme
mytheme <- theme_bw(base_size = 10)+
  theme(axis.title = element_text(size = 10, hjust = 0.5),
        plot.title.position = "panel",
        panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.25),
        plot.caption = element_text(hjust = 0),
        legend.box = "vertical",
        legend.location = "plot",
        axis.gridlines = element_line(color = "grey", linewidth = 0.25),
        axis.ticks = element_line(color = "black", linewidth = 0.5),
        axis.grid = element_blank())

#set plot theme
theme_set(mytheme)
```

# Preparation

```{r include=FALSE}
#upload test and training data for methane
methane_train <- read.csv("./Data/Processed/methane_training.csv")
methane_test <- read.csv("./Data/Processed/methane_test.csv")
methane_all <- read.csv("./Data/Processed/methane_all.csv")

#create ts object
methane_train_ts <-ts(methane_train$average, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

methane_test_ts <-ts(methane_test$average, 
                start = c(2022,1), 
                end = c(2024, 11), 
                frequency = 12)

methane_all_ts <-ts(methane_all$average, 
                start = c(1983,7), 
                end = c(2024, 11), 
                frequency = 12)

#create msts
methane_train_msts <- msts(methane_train$average, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

methane_test_msts <- msts(methane_test$average, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))

#upload test and training data for temperature
tempanom_train <- read.csv("./Data/Processed/tempanom_training.csv")
tempanom_test <- read.csv("./Data/Processed/tempanom_test.csv")
tempanom_all <- read.csv("./Data/Processed/tempanom_all.csv")

#create ts object
tempanom_train_ts <-ts(tempanom_train$temp_anomaly, 
                start = c(1983,7), 
                end = c(2021, 12), 
                frequency = 12)

tempanom_test_ts <-ts(tempanom_test$temp_anomaly, 
                start = c(2022,1), 
                end = c(2024, 11), 
                frequency = 12)

tempanom_all_ts <-ts(tempanom_all$temp_anomaly, 
                start = c(1983,7), 
                end = c(2024, 11), 
                frequency = 12)

#create msts
tempanom_train_msts <- msts(tempanom_train$temp_anomaly, 
                            seasonal.periods =c(4,12),
                            start=c(1983,7)) 

tempanom_test_msts <- msts(tempanom_test$temp_anomaly, 
                            seasonal.periods =c(4,12),
                            start=c(2022,1))
```

## ACF and PACF

```{r echo=FALSE, warning=FALSE}
#check acf and pacf
ACF_Plot <- autoplot(Acf(tempanom_train_ts, lag=40, plot = FALSE), 
                     ylim=c(-0.5,1), 
                     main="ACF plot")
PACF_Plot <- autoplot(Pacf(tempanom_train_ts, lag=40, plot = FALSE), 
                      ylim=c(-0.5,1), 
                      main="PACF plot")
plot_grid(ACF_Plot, PACF_Plot)
```

# Correlations

```{r}

methane_temp <- data.frame(methane = methane_train$average,
                           tempanom = tempanom_train$temp_anomaly)

methane_temp_cor <- cor(methane_temp)
```

# Model Fitting

```{r}
arimax_fit <- auto.arima(tempanom_train_ts,
                         xreg = methane_train_ts)
summary(arimax_fit)
checkresiduals(arimax_fit)

forecast_arimax <- forecast(arimax_fit, 
                            xreg = methane_test_ts, 
                            h = length(methane_test))
```

```{r}
fit <- nnetar(tempanom_train_msts, xreg = methane_train_msts)
forecasted <- forecast(fit, xreg = methane_test_msts)

```

Models to check out:

VAR (Vector AutoRegression) models
Random Forests, Gradient Boosting, and Neural Networks
https://stackoverflow.com/questions/51764339/how-to-predict-using-var-with-exogenous-variables-in-r 
LSTM

Seasonal Autoregressive Integrated Moving-Average with Exogenous Regressors (SARIMAX) with Exogenous Regressors (SARIMAX) is an extension of the SARIMA model that also includes the modeling of exogenous variables.

The SARIMAX method can also be used to model the subsumed models with exogenous variables, such as ARX, MAX, ARMAX, and ARIMAX.

The method is suitable for univariate time series with trend and/or seasonal components and exogenous variables.

Vector Autoregression Moving-Average with Exogenous Regressors (VARMAX) is an extension of the VARMA model that also includes the modeling of exogenous variables. It is a multivariate version of the ARMAX method.
